---
title: "Full code"
output: html_document
date: '2023-03-30'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r, warning=F, echo=FALSE}
library(tidyverse)
library(rio)
library(asbio)
library(tcltk)
library(tidyr)
library(dplyr)
library(Stat2Data)
```

```{r}
Sleep_Efficiency <- read_csv("Sleep_Efficiency.csv")
Sleep_Efficiency <- import("Sleep_Efficiency.csv")
```

Renaming variables

```{r}
sleep_dataset <- Sleep_Efficiency|>
  rename(wakeup_time = `Wakeup time`, sleep_duration = `Sleep duration`, sleep_efficiency = `Sleep efficiency`, rem_sleep_percentage = `REM sleep percentage`, caffeine_consumption = `Caffeine consumption`, age = Age, bedtime = Bedtime)
```

Selecting variables of interest

```{r}
sleep_dataset <- sleep_dataset|>
  select(wakeup_time, bedtime, sleep_duration, sleep_efficiency, rem_sleep_percentage, caffeine_consumption, age)

```

Creating age groups

```{r}
sleep_dataset <- sleep_dataset %>% 
  filter(age >= 18) %>%
  mutate(
    age_group = dplyr::case_when(
      age >= 18 & age <= 23 ~ "18-23", ##typical college students 
      age >= 24 & age <= 30 ~ "24-30", ##young working population 
      age >= 31 & age <= 40 ~ "31-40", ##middle working age  
      age >= 41 & age <= 50 ~ "41-50", ##middle working age 
      age >= 51 & age <= 60 ~ "51-60", ##middle working age 
      age >= 61 & age <= 70 ~ "61-70", #typical retirement/end of career phase 
    ),
     age_group = factor(
      age_group,
      level = c("18-23", "24-30","31-40", "41-50", "51-60", "61-70")
    )
  )
```

```{r}
sleep_dataset <- sleep_dataset %>%
    mutate(levels = cut(caffeine_consumption, breaks = c(-Inf, 0,  25, 50, 75),
        labels = c("None", "Low", "Medium", "High")))
```

```{r}
sleep_dataset <- sleep_dataset  |>
  drop_na("levels", "age_group")
```

add ggplot 

Checking assumptions

Summary statistics 

```{r}
sleep_dataset |>
  group_by(age_group) |>
  summarise(Group_mean=mean(rem_sleep_percentage), Sample_size=n(), standard_dev=sd(rem_sleep_percentage))
```


```{r}
sleep_dataset |>
  group_by(age_group) |>
  summarise(Group_mean=mean(caffeine_consumption), Sample_size=n(), standard_dev=sd(caffeine_consumption))
```

```{r, echo=FALSE}
qplot(x=age_group, y= rem_sleep_percentage, data=sleep_dataset, ylab="REM Sleep percentage", geom="boxplot") +
  theme(axis.title = element_text(size = 20))
```
Pretty symmetrical in terms of variability 
One outlier

```{r}
young <- sleep_dataset %>%
  filter(age_group == "18-23")
young_working <- sleep_dataset %>%
  filter(age_group == "24-30")
middle_working <- sleep_dataset %>%
  filter(age_group == "31-40")
middle_working1 <- sleep_dataset %>%
  filter(age_group == "31-40")
middle_working2 <- sleep_dataset %>%
  filter(age_group == "41-50")
near_retirement<- sleep_dataset %>%
  filter(age_group == "51-60")
typical_retirement <- sleep_dataset %>%
  filter(age_group == "61-70")
```

```{r}
ggplot(sleep_dataset, aes(x = levels, y = rem_sleep_percentage, color = age_group)) +
  geom_point()
```

Correlation
```{r}
cor(young$caffeine_consumption,young$rem_sleep_percentage)
```

```{r}
cor(young_working$caffeine_consumption,young_working$rem_sleep_percentage)
```
```{r}
cor(middle_working$caffeine_consumption,middle_working$rem_sleep_percentage)
```
```{r}
cor(middle_working1$caffeine_consumption,middle_working1$rem_sleep_percentage)
```
```{r}
cor(middle_working2$caffeine_consumption,middle_working2$rem_sleep_percentage)
```
```{r}
cor(near_retirement$caffeine_consumption,near_retirement$rem_sleep_percentage)
```
```{r}
cor(typical_retirement$caffeine_consumption,typical_retirement$rem_sleep_percentage)
```
ANOVA interaction model 
```{r}
aov_sleep <- aov(rem_sleep_percentage~levels*age_group, data=sleep_dataset)
summary(aov_sleep)
```

**Logistic regression model**

```{r}
model <- glm(age_group~ levels, data = sleep_dataset, family = binomial)
summary(model)
```


```{r}
# Create a contingency table of the two categorical variables
cont_table <- table(sleep_dataset$levels, sleep_dataset$age_group)

# Run a chi-squared test
chi_squared <- chisq.test(cont_table)

# Print the results of the chi-squared test
print(chi_squared)
```
P value: very small reject null - there is a difference in age group and caffeine consumption 

```{r}
with(sleep_dataset,interaction.plot(age_group, levels, rem_sleep_percentage))
```


```{r}
write.csv(sleep_dataset, "New sleep data.csv", row.names=FALSE)
```

Post ANOVA models 
```{r}
plot(aov_sleep, which=1)
```

```{r}
plot(aov_sleep, which=2)
```
